---
title: ""
format:
  revealjs:
    theme: simple
    slide-number: false
    center: false
    transition: none
    width: 1280
    height: 720
    scrollable: false
    embed-resources: true
execute:
  echo: false
  message: false
  warning: false
include-in-header:
  text: |
    <style>
      .reveal .slides{ text-align:left; }
      .reveal section{ overflow: hidden; }

      .topbar{
        background:#2FA892;
        color:#ffffff;
        font-size:44px;
        font-weight:800;
        padding:18px 28px;
        border-radius:0px;
        margin: 0 0 14px 0;
      }

      .section-title{
        font-size:18px;
        font-weight:900;
        letter-spacing:0.2px;
        margin: 0 0 8px 0;
        text-transform: uppercase;
      }

      .body-text{
        font-size:16px;
        line-height:1.25;
        color:#2b2b2b;
        margin-bottom: 10px;
      }

      .hl-orange{ color:#F07D2B; font-weight:900; }

      .bottombar{
        margin-top: 14px;
        background:#EDEDED;
        color:#3a3a3a;
        padding: 12px 18px;
        font-size: 20px;
        font-weight: 800;
      }

      .reveal p{ margin: 0.2em 0; }
    </style>
---

```{r}
#| include: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tibble)

has_ggrepel <- requireNamespace("ggrepel", quietly = TRUE)
```

## {.custom-slide}

<div class="topbar">More funding needed to complete projects on time</div>

::: columns
::: {.column width="50%"}
<div class="section-title">Actual resources higher than plan</div>

<div class="body-text">
The total workforce allocation is <span class="hl-orange">1,026 FTE above the expected plan</span> as of the end of April 2024.
The actual FTEs are 44% higher than expected, and an average is expected to continue for the remainder of the year.
</div>

```{r}
#| include: false
# -----------------------------
# DATAFRAME FAKE (WORKFORCE: ACTUAL vs PLAN)
# OBS: para ficar como o exemplo do slide, o "ACTUAL" só existe até APR-24.
# Depois disso, só mantemos o PLAN (linha pontilhada).
# -----------------------------
months_1 <- c("Sep-23","Oct-23","Nov-23","Dec-23","Jan-24","Feb-24","Mar-24","Apr-24","May-24","Jun-24","Jul-24","Aug-24","Sep-24")

plan <- c(120, 450, 410, 360, 310, 270, 200, 150, 150, 150, 150, 150, 150)

# Diferenças escolhidas para somar 1,026 no período Sep-23..Apr-24
diff_8 <- c(180, 80, 95, 110, 130, 150, 150, 131)  # soma = 1,026

actual_8 <- plan[1:8] + diff_8

df_workforce <- tibble(
  month = factor(months_1, levels = months_1),
  plan = plan,
  actual = c(actual_8, rep(NA_real_, length(months_1) - 8))  # <- ACTUAL só até Apr
)

df_actual <- df_workforce |> filter(!is.na(actual)) |>
  mutate(ymin = pmin(plan, actual), ymax = pmax(plan, actual))  # ribbon seguro

total_overplan_8 <- df_workforce |>
  mutate(over_plan = actual - plan) |>
  filter(month %in% months_1[1:8]) |>
  summarise(total = sum(over_plan, na.rm = TRUE)) |>
  pull(total)
```

```{r}
#| fig-width: 6.2
#| fig-height: 3.3
#| dpi: 180
#| out-width: "100%"
# -----------------------------
# PLOT 1: Workforce actual vs plan by month
# AJUSTE IMPORTANTE:
# - o laranja é SOMENTE a área "acima do plano" (entre PLAN e ACTUAL),
#   não uma área até o zero.
# -----------------------------
p1 <- ggplot() +
  # Área do overrun (entre plan e actual)
  geom_ribbon(
    data = df_actual,
    aes(x = month, ymin = ymin, ymax = ymax, group = 1),
    fill = "#F07D2B", alpha = 0.70
  ) +
  # Linha ACTUAL (só até Apr-24)
  geom_line(
    data = df_actual,
    aes(x = month, y = actual, group = 1),
    color = "#D85A00", linewidth = 1.8
  ) +
  # Linha PLAN (até o fim)
  geom_line(
    data = df_workforce,
    aes(x = month, y = plan, group = 1),
    color = "#7A7A7A", linewidth = 1.3, linetype = "dashed"
  ) +
  scale_y_continuous(
    limits = c(0, 600),
    breaks = seq(0, 600, 50)
  ) +
  labs(
    title = "Workforce: actual versus plan by month",
    x = NULL, y = "# OF FTE"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    axis.text.x = element_text(size = 9),
    panel.grid.minor = element_blank(),
    plot.margin = margin(6, 10, 6, 6)
  )

# Anotação principal
p1 <- p1 +
  annotate(
    "text",
    x = "Apr-24", y = 260,
    label = paste0("ACTUAL SEP'23 to APR'24\n", format(total_overplan_8, big.mark=","), " FTE above plan"),
    color = "#D85A00",
    fontface = "bold",
    hjust = 0, vjust = 1,
    size = 3.3
  ) +
  annotate(
    "text",
    x = "Sep-24", y = 135,
    label = "PLAN",
    color = "#7A7A7A",
    fontface = "bold",
    hjust = 1,
    size = 3.0
  ) +
  coord_cartesian(clip = "off")

p1
```

:::
::: {.column width="50%"}
<div class="section-title">Ascent project causing 84% of overrun</div>

<div class="body-text">
Most projects were close to plan expectations (within 20 FTE of the monthly estimate).
However, the <b>Ascent project is markedly over plan</b> due to a major scope change requiring additional workforce support.
</div>

```{r}
#| include: false
# -----------------------------
# DATAFRAME FAKE (FTE over plan by project)
# -----------------------------
months_2 <- c("Sep-23","Oct-23","Nov-23","Dec-23","Jan-24","Feb-24")

df_proj <- tribble(
  ~project,   ~month,   ~fte_over_plan,
  "ASCENT",   "Sep-23", 190,
  "ASCENT",   "Oct-23",  50,
  "ASCENT",   "Nov-23",  70,
  "ASCENT",   "Dec-23", 110,
  "ASCENT",   "Jan-24", 140,
  "ASCENT",   "Feb-24", 145,

  "CATALYST", "Sep-23",  12,
  "CATALYST", "Oct-23",  14,
  "CATALYST", "Nov-23",  16,
  "CATALYST", "Dec-23",  12,
  "CATALYST", "Jan-24",  14,
  "CATALYST", "Feb-24",  10,

  "ODYSSEY",  "Sep-23",  -5,
  "ODYSSEY",  "Oct-23",  -2,
  "ODYSSEY",  "Nov-23",   0,
  "ODYSSEY",  "Dec-23",  -1,
  "ODYSSEY",  "Jan-24",  -3,
  "ODYSSEY",  "Feb-24",  -6
) |>
  mutate(
    month = factor(month, levels = months_2),
    project = factor(project, levels = c("ASCENT","CATALYST","ODYSSEY"))
  )

df_last <- df_proj |>
  group_by(project) |>
  filter(month == last(month)) |>
  ungroup()

col_map <- c("ASCENT"="#000000", "CATALYST"="#9A9A9A", "ODYSSEY"="#B8B8B8")
lw_map  <- c("ASCENT"=2.4, "CATALYST"=1.4, "ODYSSEY"=1.4)
```

```{r}
#| fig-width: 6.2
#| fig-height: 3.3
#| dpi: 180
#| out-width: "100%"
# -----------------------------
# PLOT 2: FTE over plan by project
# -----------------------------
p2 <- ggplot(df_proj, aes(x = month, y = fte_over_plan, group = project, color = project)) +
  geom_hline(yintercept = 0, color = "#CFCFCF", linewidth = 0.8) +
  geom_line(aes(linewidth = project), alpha = 0.98) +
  scale_color_manual(values = col_map) +
  scale_linewidth_manual(values = lw_map) +
  scale_y_continuous(
    limits = c(-20, 200),
    breaks = seq(-20, 200, 20)
  ) +
  labs(
    title = "Number of FTE over plan by project",
    x = NULL, y = "# FTE OVER PLAN"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 12),
    axis.text.x = element_text(size = 9),
    panel.grid.minor = element_blank(),
    plot.margin = margin(6, 6, 6, 6)
  )

# Labels no final das linhas (direita)
if (has_ggrepel) {
  p2 <- p2 +
    ggrepel::geom_text_repel(
      data = df_last,
      aes(label = project),
      nudge_x = 0.35,
      direction = "y",
      hjust = 0,
      size = 3.0,
      segment.size = 0.2,
      min.segment.length = 0,
      box.padding = 0.2
    ) +
    coord_cartesian(xlim = c(1, length(months_2) + 0.8), clip = "off")
} else {
  p2 <- p2 +
    geom_text(
      data = df_last,
      aes(label = project),
      nudge_x = 0.25,
      hjust = 0,
      size = 3.0,
      show.legend = FALSE
    ) +
    coord_cartesian(xlim = c(1, length(months_2) + 0.8), clip = "off")
}

p2
```

:::
:::

<div class="bottombar">Approve funding to cover the cost of additional resources and ensure projects are completed on time.</div>
