---
title: ""
format:
  revealjs:
    theme: simple
    slide-number: false
    center: false
    transition: none
    width: 1280
    height: 720
    scrollable: false
    embed-resources: true
execute:
  echo: false
  message: false
  warning: false
include-in-header:
  text: |
    <style>
      .topbar{
        background:#0B4F7A;
        color:#ffffff;
        font-size:40px;
        font-weight:700;
        padding:18px 28px;
        border-radius:6px;
        margin-bottom:18px;
      }
      .section-title{
        font-size:18px;
        font-weight:800;
        letter-spacing:0.2px;
        margin: 0 0 8px 0;
        text-transform: uppercase;
      }
      .body-text{
        font-size:16px;
        line-height:1.25;
        color:#2b2b2b;
      }
      .hl-blue{ color:#1f77b4; font-weight:700; }
      .hl-orange{ color:#ff7f0e; font-weight:700; }
      .recommendation{
        margin-top:10px;
        font-size:16px;
      }
      .recommendation b{
        font-weight:900;
      }
      .panel{
        padding: 6px 10px;
      }
      .reveal .slides{
        text-align:left;
      }
      .reveal section{
        overflow: hidden;
      }
    </style>
---

```{r}
#| include: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tibble)

has_ggrepel <- requireNamespace("ggrepel", quietly = TRUE)
```

## {.custom-slide}

<div class="topbar">Room bookings up but some areas of concern</div>

::: columns
::: {.column width="50%"}
<div class="panel">

<div class="section-title">Regional split to be reviewed</div>

<div class="body-text">
Room bookings for hotels <span class="hl-blue">based in the south</span> have increased compared with 2019.
It is the opposite for our <span class="hl-orange">northern hotels</span> who are still struggling to recover bookings following the pandemic.
</div>

<div class="recommendation body-text">
<b>RECOMMENDATION:</b> Understand and share best practices from <span class="hl-blue">south based</span> hotels to increase room bookings in the <span class="hl-orange">north</span>.
</div>

```{r}
#| include: false
df_city_change <- tribble(
  ~city,         ~region, ~pct_change_2022,
  "Cardiff",     "South",  0.45,
  "Oxford",      "South",  0.22,
  "London",      "South",  0.18,
  "Portsmouth",  "South",  0.14,
  "Bristol",     "South",  0.08,
  "Newcastle",   "North", -0.10,
  "Leeds",       "North", -0.16,
  "Liverpool",   "North", -0.21,
  "York",        "North", -0.28,
  "Manchester",  "North", -0.42
)

df_line <- df_city_change |>
  expand_grid(year = c(2019, 2022)) |>
  mutate(pct_change = ifelse(year == 2019, 0, pct_change_2022)) |>
  arrange(city, year)

pal_region <- c("South" = "#1f77b4", "North" = "#ff7f0e")
```

```{r}
#| fig-width: 6
#| fig-height: 3.2
#| dpi: 160
#| out-width: "100%"
p1 <- ggplot(df_line, aes(x = year, y = pct_change, group = city, color = region)) +
  geom_hline(yintercept = 0, linewidth = 0.4, alpha = 0.6) +
  geom_line(linewidth = 1.2, alpha = 0.95) +
  geom_point(size = 2) +
  scale_color_manual(values = pal_region) +
  scale_x_continuous(breaks = c(2019, 2022), limits = c(2019, 2022.6)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(-0.55, 0.55)) +
  labs(
    title = "Hotel room booking comparison between 2019 and 2022",
    subtitle = "south region vs north region",
    x = NULL, y = "Change vs 2019"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(size = 11),
    panel.grid.minor = element_blank()
  )

df_end <- df_line |> filter(year == 2022)

if (has_ggrepel) {
  p1 <- p1 +
    ggrepel::geom_label_repel(
      data = df_end,
      aes(label = city),
      size = 3.2,
      label.size = 0,
      fill = "white",
      alpha = 0.9,
      box.padding = 0.25,
      point.padding = 0.20,
      nudge_x = 0.35,
      direction = "y",
      segment.size = 0.2,
      min.segment.length = 0,
      seed = 7
    )
} else {
  p1 <- p1 +
    geom_label(
      data = df_end,
      aes(label = city),
      size = 3.0,
      label.size = 0,
      fill = "white",
      alpha = 0.9,
      hjust = 0,
      nudge_x = 0.05,
      show.legend = FALSE
    )
}

p1
```

</div>
:::

::: {.column width="50%"}
<div class="panel">

<div class="section-title">Reduction of single rooms in London area</div>

<div class="body-text">
<span class="hl-blue">Single room bookings in London</span> have reduced by <b>81%</b> since 2019.
This is due to the increased amount of hybrid working, reduced time in the office with the need to stay overnight.
</div>

<div class="recommendation body-text">
<b>RECOMMENDATION:</b> Explore the creation of an incentive scheme to reward business travellers back to London.
</div>

```{r}
#| include: false
single_2019 <- 64
single_2022 <- round(single_2019 * (1 - 0.81))

df_london <- tribble(
  ~year, ~room_type,     ~bookings,
  2019,  "Single room",  single_2019,
  2019,  "Double room",  180,
  2019,  "Twin room",    55,
  2019,  "Family room",  40,
  2019,  "Suite",        15,
  2022,  "Single room",  single_2022,
  2022,  "Double room",  210,
  2022,  "Twin room",    60,
  2022,  "Family room",  45,
  2022,  "Suite",        17
)

fill_map <- c(
  "Single room" = "#1f77b4",
  "Double room" = "grey75",
  "Twin room"   = "grey65",
  "Family room" = "grey55",
  "Suite"       = "grey45"
)

df_london <- df_london |>
  mutate(room_type = factor(room_type, levels = c("Single room","Double room","Twin room","Family room","Suite")))
```

```{r}
#| fig-width: 6
#| fig-height: 3.2
#| dpi: 160
#| out-width: "100%"
p2 <- ggplot(df_london, aes(x = factor(year), y = bookings, fill = room_type)) +
  geom_col(width = 0.55) +
  scale_fill_manual(values = fill_map) +
  labs(
    title = "Room booking comparison: London",
    x = NULL, y = "Bookings", fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold", size = 13),
    panel.grid.minor = element_blank()
  )

p2
```

</div>
:::
:::
